---
title: Telco customer churn
author:
  - name: Ketao Li
    affiliation: York University
    email:  liketao@yahoo.com
  - name: Kush Halani
    affiliation: York University
    email:  kush.halani@ontariotechu.net
  - name: Josue Romain
    affiliation: York University
    email:  josue.rolland.romain@gmail.com    
  - name: Juan Peña
    affiliation: York University
    email:  jppena62@my.yorku.ca
  - name: Priyanka Patil
    affiliation: York University
    email:  priyanka181994@gmail.com    
abstract: >
  An abstract of less than 150 words.

output:
  rticles::rjournal_article:
    includes:
      in_header: preamble.tex
---


```{r echo=FALSE, message=FALSE, warnings=FALSE}
# load required libraries
library(ggplot2) # plotting lib
library(gridExtra) # arrange grids
library(dplyr)  # data manipuation
library(mice)  # data imputing
library(corrplot) # correlation matrix plotting/printing
library(pROC) # to measure model performance
library(leaflet) # maps
library(RColorBrewer) # color palettes
library(VIM) # missing value analysis 
library(lattice) # another data plotting library
library(mapview) # saves map objects as file
library(png) # deals with png file measurements
library(knitr) #
library(party) # classification tree
library(klaR) # naive bayes
library(xtable) # tabular data formatting 
library(caret) # predictive models

# Clean all variables that might be left by other script to avoid collusion
rm(list=ls(all=TRUE))
# set xtable properties for the project
options(xtable.floating = TRUE)
options(xtable.timestamp = "")
options(xtable.comment = FALSE)

# pick palettes
mainPalette = brewer.pal(8,"Dark2")
# set a sample size
SampleSize = 3000
```

```{r global_options, include=FALSE}
# make the images flow nicely
knitr::opts_chunk$set(fig.pos = 'H')
```


## Background

In an industry as competitive as Telecom, leading companies know that the key to success is not just about acquiring new customers, but rather, retaining existing ones. But how do you know which customers are at risk and why, and which negative experiences and interactions have the biggest impact on churn across touchpoints and channels over time


## Objective

The objective of this research is to find a supervised, binary classification model that would provide accurate forecast of telco customer churn.

# Data Analysis

The data set we are going to use for our research contains  customer’s attributes. There are over 7044 records. It has been sourced from [Kaggle](https://www.kaggle.com/blastchar/telco-customer-churn)


## Data Dictionary



Column Name            | Column Description  
-----------------------| ------------------- 
customerID             | Customer ID 
gender                 | Whether the customer is a male or a female
SeniorCitizen          | Whether the customer is a senior citizen or not (1, 0)
Partner                | Whether the customer has a partner or not (Yes, No)
Dependents             | Whether the customer has dependents or not (Yes, No)
tenure                 | Number of months the customer has stayed with the company
PhoneService           | Whether the customer has a phone service or not (Yes, No)
MultipleLines          | Whether the customer has multiple lines or not (Yes, No, No phone service)
InternetService        | Customer’s internet service provider (DSL, Fiber optic, No)
OnlineSecurity         | Whether the customer has online security or not (Yes, No, No internet service)
OnlineBackup           | Whether the customer has online backup or not (Yes, No, No internet service)
DeviceProtection       | Whether the customer has device protection or not (Yes, No, No internet service)
TechSupport            | Whether the customer has tech support or not (Yes, No, No internet service)
StreamingTV            | Whether the customer has streaming TV or not (Yes, No, No internet service)
StreamingMovies        | Whether the customer has streaming movies or not (Yes, No, No internet service)
Contract               | The contract term of the customer (Month-to-month, One year, Two year)
PaperlessBilling       | Whether the customer has paperless billing or not (Yes, No)
PaymentMethod          | The customer’s payment method (Electronic check, Mailed check, Bank transfer (automatic), Credit card (automatic))
MonthlyCharges         | The amount charged to the customer monthly
TotalCharges           | The total amount charged to the customer
Churn                  | Whether the customer churned or not (Yes or No)


## Data Exploration

Let's take a close look at the data set.

```{r message=FALSE, warning=FALSE}
customerData = read.csv("../data/WA_Fn-UseC_-Telco-Customer-Churn.csv", header = TRUE, na.strings = c("NA","","#NA"),sep=",")

```
  

To have the full picture of the data let's print the data summary and sample.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results='asis'}
print(xtable(summary(customerData[,1:8])), include.rownames = FALSE, scalebox=.7)
print(xtable(summary(customerData[,9:16])), include.rownames = FALSE, scalebox=.7)
print(xtable(summary(customerData[,17:21]), caption = "\\tt telco customer churn data Summary", 
             label = "data_head"), include.rownames = FALSE, scalebox=.7)
```
\newpage
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results='asis'}
print(xtable(customerData[1:10,1:12]), scalebox=.6)
print (xtable(customerData[1:10,13:21],
  caption = "\\tt telco customer churn data", label = "data_head"), include.rownames = F,
  scalebox = .6)
```

```{r}
#DATA EXPLORATION

#To see the names of the rows in the dataset
names(customerData)

#Display the dataset structure and sumary
str(customerData)

#Display first rows of the dataset
head(customerData)

#To select just the continuous variables and summarise it
library(dplyr)
continues <- select_if(customerData, is.numeric)
#Sumarize the variables to find NA's and outliers
summary(continues)

#Display the factor columns and summarise it
factorColumns <- select_if(customerData, is.factor)
summary(factorColumns)
```

#Make a chart for each factor type column.

```{r test1,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test1 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=gender)) +
                        geom_bar()

```

```{r test2,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test2 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=Partner)) +
        geom_bar()
```

```{r test3,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test3 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=Dependents)) +
        geom_bar()
```

```{r test4,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test4 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=PhoneService)) +
        geom_bar()
```

```{r test5,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test5 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=MultipleLines)) +
        geom_bar()
```

```{r test6,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test6 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=InternetService)) +
        geom_bar()
```

```{r test7,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test7 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=OnlineSecurity)) +
        geom_bar()
```



```{r test8,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test8 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=OnlineBackup)) +
        geom_bar()
```


```{r test9,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test9 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=DeviceProtection)) +
        geom_bar()
```


```{r test10,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test10 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=TechSupport)) +
        geom_bar()
```


```{r test11,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test11 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=StreamingTV)) +
        geom_bar()
```


```{r test12,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test12 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=StreamingMovies)) +
        geom_bar()
```

```{r test13,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test13 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=Contract)) +
        geom_bar()
```


```{r test14,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test14 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=PaperlessBilling)) +
        geom_bar()
```


```{r test15,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test15 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=PaymentMethod)) +
        geom_bar()
```


```{r test16,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test16 ", out.width="1.1\\linewidth"}
ggplot(factorColumns, aes(x=Churn)) +
        geom_bar()
```

#As we can see in the continues' summary, there are 11 NA's in the Totalcharges column and some
#outliers in Monthlycharges and TotalCharges.
# Let's take a look of those columns in a graph to better understanding

```{r test17,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test17 ", out.width="1.1\\linewidth"}
ggplot(continues, aes(x=MonthlyCharges)) +
        geom_density(alpha= .2, fill="#FF6666")
```

#There is no normal distribution in Monthlycharges

#Let's see now Totalchrges
```{r test18,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Test18 ", out.width="1.1\\linewidth"}
ggplot(continues, aes(x=TotalCharges)) +
        geom_density(alpha= .2, fill="#FF6666")
```

#Can we make a hypothesis about the relation of charges total and monthly with the client churn?
```



## Summary

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).

\bibliography{RJreferences}
